---
title: "NSW Cancer Institute pre-interview questions"
author: "David Kennedy"
date: '2022-11-08'
output: html_document
---

```{r setup}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.asp = 0.8,
	fig.height = 12,
	fig.retina = 2,
	fig.width = 12,
	warning = FALSE,
	message=FALSE,
	dev = "jpeg",
	dpi = 600,
	units="cm",
	out.width = "100%"
)
pacman::p_load(tidyverse,here,janitor,flextable, lubridate, survival, magrittr,patchwork,sqldf)

```

# Question 1

### Use SQL code to update the missing patient's first name and last name in Patient table from stay table.

```{r question-1}
raw <- here('raw_data.xlsx')
patient <- readxl::read_xlsx(raw, sheet='patient')
stay <- readxl::read_xlsx(raw, sheet='stay')

sqldf('SELECT patient.mrn, 
coalesce(patient."First Name", stay."First Name") AS "First Name",
coalesce(patient."Last Name", stay."Last Name") AS "Last Name",
patient.DOB, 
patient.Gender
      FROM patient  
      LEFT JOIN stay
      ON patient.mrn = stay.mrn')

```

# Question 2

### Load Simulacrum data

```{r load-data}
sim_av_patient <- read_csv(list.files(here('data'), 
                                      pattern='sim_av_patient.csv', 
                                      recursive=T, 
                                      full.names=T))

sim_av_tumour <- read_csv(list.files(here('data'), 
                                     pattern='sim_av_tumour.csv', 
                                     recursive=T, 
                                     full.names=T))
```

### 2a. What is the sex distribution of patients?

```{r question-2a}

(sex_distribution <- sim_av_patient %>% mutate(sex_factor=factor(SEX, levels=c(1,2,9), 
                                                                 labels=c('Males', 'Females', 'Unknown'))) %>% 
    tabyl(sex_factor) %>% 
    adorn_pct_formatting(digits=1) %>% 
    flextable() %>%  
    autofit()  %>% 
    set_header_labels(.,
                      sex_factor = "Sex",
                      n = "Count", 
                      percent = "%")
)

```

### 2b. Prepare a dataset with all lung cancer patients (SITE_ICD10_O2_3CHAR= C34) and their cause of deaths.

```{r}

(lung_cancer_patients_cod <-
  sim_av_patient %>% select(PATIENTID, contains('DEATHCAUSE')) %>%
  right_join(
    sim_av_tumour %>% filter(SITE_ICD10_O2_3CHAR == 'C34') %>%
      select(PATIENTID, SITE_ICD10_O2_3CHAR),
    by = c('PATIENTID')
  )
)
```

```{r}
(lung_cancer_patients_2013<-
   sim_av_patient %>% select(PATIENTID, NEWVITALSTATUS, VITALSTATUSDATE) %>%
   right_join(
     sim_av_tumour %>% filter(SITE_ICD10_O2_3CHAR %in% c('C34')) %>%
       mutate(year=lubridate::year(DIAGNOSISDATEBEST)) %>% 
       filter(year==2013) %>% 
       select(PATIENTID, SITE_ICD10_O2_3CHAR, DIAGNOSISDATEBEST),
     by = c('PATIENTID')
   ) %>% 
   mutate(
     os_days = as.duration(DIAGNOSISDATEBEST %--% VITALSTATUSDATE) / ddays(1),
     flag=case_when(NEWVITALSTATUS=='D' ~ 1, 
                    TRUE ~ 0)
   )
)

print(survfit(Surv(os_days, flag) ~ 1, data = lung_cancer_patients_2013), print.rmean=TRUE)

km_fit <- survfit(Surv(os_days, flag) ~ 1, data = lung_cancer_patients_2013)
plot(km_fit, xlab="Days", main = 'K-M Plot') 
```

# Question 3

I do not have access to Power BI, so have used R to investigate the ethnicity variable.

```{r data-processing}
sim_av_patient$ETHNICITY_factor =  recode_factor(
  sim_av_patient$ETHNICITY,
  "A" = "(White) British",
  "B" = "(White) Irish",
  "C" = "Any other White background",
  "D" = "White and Black Caribbean",
  "E" = "White and Black African",
  "F" = "White and Asian",
  "G" = "Any other mixed background",
  "H" = "Indian",
  "J" = "Pakistani",
  "K" = "Bangladeshi",
  "L" = "Any other Asian background",
  "M" = "Caribbean",
  "N" = "African",
  "P" = "Any other Black background",
  "R" = "Chinese",
  "S" = "Any other ethnic group" ,
  "Z" = "Not stated",
  "X" = "Not known"
)

sim_av_patient$ETHNICITY_factor %<>% forcats::fct_explicit_na()

ethnicity_data <- sim_av_patient %>%
  mutate(british = case_when(ETHNICITY_factor == "(White) British" ~ 1,
                             TRUE ~ 0)) %>%
  filter(british == 0) %>% 
  mutate(known_ethnicity = case_when(
    ETHNICITY_factor %in% c('(Missing)', 'Not stated', 
                            'Not known', '0', '8', 'CA') ~ 'Unknown ethnicity',
    TRUE ~ 'Known ethnicity'
  )) 

mycolors <- c("#002664", "#D7153A")
```

```{r column-gph}
column_gph <- ethnicity_data %>% 
  count(known_ethnicity, sort = TRUE) %>%
  mutate(prop = n / sum(n)) %>% 
  ggplot(aes(x = known_ethnicity, y = prop, fill=factor(known_ethnicity))) +
  geom_col() +
    geom_text(aes(label = round(prop*100,1)), color = "white", size = 10, position = position_stack(vjust = 0.5)) + 
    scale_fill_manual(values=mycolors) +
  theme_bw() +
    scale_y_continuous(labels = scales::percent, breaks=seq(from = 0, to = 1, by = .1)) +
    scale_x_discrete( expand = c(0, 0)) +
    theme(legend.position='none',
      axis.line = element_blank(),
      axis.ticks = element_blank(),
      strip.text = element_text(face = "bold"),
      panel.grid = element_blank(),
      panel.spacing = unit(2, "lines"),
      text = element_text(size = 10)
    ) +
  labs(x='', y='', title='Most records have unknown ethnicity data') 

```

```{r point-gph}
point_gph <- ethnicity_data %>%
  count(known_ethnicity,ETHNICITY_factor, sort = TRUE) %>%
  mutate(prop = n / sum(n)) %>% 
  mutate(ETHNICITY_factor = fct_reorder(ETHNICITY_factor, (n))) %>%
    filter(known_ethnicity=='Known ethnicity') %>% 
  ggplot(aes(x = ETHNICITY_factor, y = prop, group=1)) +
  geom_point(size=8, aes(col=known_ethnicity)) +
    scale_color_manual(values=mycolors) +
  labs(x='', y='', title = 'The known ethnicity group is made up of 15 ethnicity classifications') +
  coord_flip() +
  theme_bw() +
  scale_y_continuous(labels = scales::percent) +
    theme(legend.position='none',
      axis.line = element_blank(),
      strip.text = element_text(face = "bold"),
      strip.background = element_rect(fill = "#CBEDFD", colour = "black", size = 1),
      panel.grid = element_blank(),
      panel.spacing = unit(2, "lines"),
      text = element_text(size = 10)
    ) +
  facet_wrap(~ (known_ethnicity), scales='free_y', ncol=1)

```

```{r missingness-gph}
missingness_overtime <- ethnicity_data %>% 
  mutate(year=lubridate::year(VITALSTATUSDATE)) %>% 
  filter(!is.na(year)) %>% 
  count(year, known_ethnicity) %>% group_by(year) %>% 
  mutate(missingness=n/sum(n)) %>% 
    filter(known_ethnicity=='Unknown ethnicity') %>% 
    ggplot(aes(x=year, y=missingness)) +
    geom_line(color='#D7153A',size=2) +
    theme_bw() +
    labs(x='Year', title='The % of records with unknown ethnicity data has increased since 2013', y='') +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(breaks=c(2013:2019)) +
    theme(
      axis.line = element_blank(),
      strip.text = element_text(face = "bold"),
      strip.background = element_rect(fill = "#CBEDFD", colour = "black", size = 1),
      panel.grid = element_blank(),
      panel.spacing = unit(2, "lines"),
      text = element_text(size = 10)
    )

  
```

```{r additional-data-processing}
age_ethnicity_data <- ethnicity_data %>% select(known_ethnicity, PATIENTID) %>% 
  left_join(sim_av_tumour %>% select(PATIENTID,AGE), by='PATIENTID') %>% 
  mutate(age_group=cut(AGE, breaks=c(-1,19,39,59,79,Inf), labels=c('0-19', '20-39', '40-59', '60-79', '80+'))) %>% 
  count(known_ethnicity,age_group) %>% 
  group_by(age_group) %>% mutate(sum=sum(n)) %>%   
  mutate(prop = map2(n, sum, ~ prop.test(.x, .y, conf.level=0.95) %>% 
                       broom::tidy())) %>%
  unnest(prop) %>% 
  ungroup()
```

```{r age-ethnicity}
age_ethnicity_gph <- 
    ggplot(data=age_ethnicity_data,aes(x=age_group, y=estimate)) +
    geom_col(aes(fill=factor(known_ethnicity)), position = position_dodge2())+
    theme_bw() +
    scale_fill_manual(values=mycolors) +
    scale_y_continuous(labels = scales::percent, breaks=seq(from = 0, to = 1, by = .1)) +
    labs(x='Age group', title='Completeness of ethnicity data varies by age group', y='', fill='') +
    theme(legend.position = c(0.87, 0.9),legend.key.size = unit(0.5, 'cm'),legend.text = element_text(size=8),
          axis.line = element_blank(),
          strip.text = element_text(face = "bold"),
          strip.background = element_rect(fill = "#CBEDFD", colour = "black", size = 1),
          panel.grid = element_blank(),
          panel.spacing = unit(2, "lines"),
          text = element_text(size = 10))

```

```{r combine-plots}
patchwork3 <- (column_gph + point_gph) / (missingness_overtime + age_ethnicity_gph)

patchwork3 + plot_annotation(
  title = 'Investigating the completeness of ethnicity data in a simulated cancer registry',
  caption = 'Data source: Simulacrum',
  theme = theme(plot.title = element_text(size = 18))) 
```
